---
title: "Inequality PSet 1 - Question #6 (code part)"
author: "JDu"
date: "`r lubridate::today()`"
output: pdf_document
---

## Load necessary libraries
```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(lfe)
library(stats)
#library(tsibble)
library(ivreg)
library(broom)
library(stargazer)

theme_set(theme_minimal())
```

## could write 6c in .Rmd - but should do 6a & 6b on paper

# Question 6d

## Question 6d(i): Construct the outcome variables

```{r}
q1 <- read_csv("./DataExercise_IV/pset_iv_data.csv") %>%
  rename(pop = Pop, earnPOW = EarnPOW, man = Manufact)

# 6d(i).A: transform var by taking nat logs
q1 <- q1 %>%
  mutate(lnpay = log(pay),
         lnpop = log(pop),
         lnearnPOW = log(earnPOW), # county earnings
         lndSSI = log(dSSI)) # SSI payments 

# 6d(i).B: construct log diff
q1 <- q1 %>%
  group_by(fips) %>% #DELETE HERE IF NEEDED
  mutate(lagpay = lag(lnpay, n = 1),
         lagpop = lag(lnpop, n = 1),
         lagearnPOW = lag(lnearnPOW, n = 1),
         lagdSSI = lag(lndSSI, n = 1)) %>%
  mutate(diff_pay = lnpay - lagpay,
         diff_pop = lnpop - lagpop,
         diff_earnPOW = lnearnPOW - lagearnPOW,
         diff_dSSI = lndSSI - lagdSSI) 
  
# 6d(i).C: construct var that is fraction of county earnings from manufacturing in 1969
q1 <- 
  q1 %>%
  filter(year == 1969) %>%
  mutate(frac_earn = man/earnPOW) %>%
  select(frac_earn, fips) %>%
  left_join(q1, by = "fips")
```

## Question 6d(ii): Construct the instruments

```{r}
# construct coal price
q1 <- q1 %>%
  mutate(coalprice = pcoalyr/pyr)

# construct change in coal reserves
q1 <- q1 %>%
  group_by(fips) %>%
  mutate(ln_coalprice = log(coalprice),
         lag_coalprice = lag(ln_coalprice, n = 1),
         diff_coalprice = ln_coalprice - lag_coalprice) %>%
  mutate(ln_coalres = log(coalres)) %>%
  mutate(valchange_cr = 
           if_else(coalres > 0,
                   diff_coalprice * ln_coalres, 0)) %>%
  mutate(valchange_crlag1 = lag(valchange_cr, n = 1),
         valchange_crlag2 = lag(valchange_cr, n = 2))

#doesn't make a diff when you do if_else - most likely is 1969?
# NOPE DOESN'T MAKE A DIFF
q1_try <- q1 %>%
  group_by(fips) %>%
  mutate(ln_coalprice = log(coalprice),
         lag_coalprice = lag(ln_coalprice, n = 1),
         diff_coalprice = ln_coalprice - lag_coalprice) %>%
  mutate(ln_coalres = if_else(coalres > 0,
                   log(coalres), 0)) %>%
  mutate(valchange_cr = diff_coalprice * ln_coalres, 0) 

q1 %>%
  filter(valchange_cr < 0)
```


## Question 6d(iii).A: OLS structural eqns

```{r}
library(moderndive) #https://cran.r-project.org/web/packages/moderndive/vignettes/why-moderndive.html
library(lfe) #felm - multiple group fixed effs? - 
#https://cran.r-project.org/web/packages/lfe/lfe.pdf (pg. 19)
#library(robustbase) 
#use lmrob just like lm - https://www.brodrigues.co/blog/2018-07-08-rob_stderr/
library(lmtest)
library(sandwich)
# https://www.r-econometrics.com/methods/hcrobusterrors/

ols_ssi_c <- lm(diff_dSSI ~ diff_earnPOW + as.factor(year)*as.factor(state) + msa + lnpop + diff_pop + frac_earn, data = q1)

ols_ssi_nc <- lm(diff_dSSI ~ diff_earnPOW + as.factor(year)*as.factor(state), data = q1)

ols_di_c <- lm(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state) + msa + lnpop + diff_pop + frac_earn, data = q1)

ols_di_nc <- lm(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state), data = q1)

# find estimates & robust SEs
coeftest(ols_ssi_c, vcov = vcovHC(ols_ssi_c, type = "HC1")) %>%
  broom::tidy() 

coeftest(ols_ssi_nc, vcov = vcovHC(ols_ssi_nc, type = "HC1")) %>%
  tidy()

coeftest(ols_di_c, vcov = vcovHC(ols_di_c, type = "HC1")) %>%
  tidy()

coeftest(ols_di_nc, vcov = vcovHC(ols_di_nc, type = "HC1")) %>%
  tidy()

```

NEED TO INTERPRET MAIN ESTIMATES (that's just the estimated coeff for diff_earnPOW aka log diff in county earnings, right?)

## Question 6d(iii).B: 1st-stage estimates

```{r}
first_c <- lm(diff_earnPOW ~ as.factor(state)*as.factor(year) + valchange_cr + msa + lnpop + diff_pop + frac_earn, data = q1)

first_nc <- lm(diff_earnPOW ~ as.factor(state)*as.factor(year) + valchange_cr, data = q1)

get_regression_table(first_c) #normal SE (not robust)
get_regression_table(first_nc) #normal SE (not robust)
```

interpret the main estimates

## Question 6d(iii).C: 2SLS structural eqn

```{r}
#https://cran.r-project.org/web/packages/ivreg/ivreg.pdf
#https://www.econometrics-with-r.org/12-2-TGIVRM.html

iv_ssi_c <- ivreg(diff_dSSI ~ 
                    diff_earnPOW + as.factor(year)*as.factor(state) + msa + 
                    lnpop + diff_pop + frac_earn 
                  | valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                    as.factor(year)*as.factor(state) + msa + 
                    lnpop + diff_pop + frac_earn, data = q1)


iv_ssi_nc <- ivreg(diff_dSSI ~ diff_earnPOW + as.factor(year)*as.factor(state) 
                   | valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                     as.factor(year)*as.factor(state), data = q1)

iv_di_c <- ivreg(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state) + 
                   msa + lnpop + diff_pop + frac_earn 
                 | valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                   as.factor(year)*as.factor(state) + msa + 
                   lnpop + diff_pop + frac_earn, data = q1)

iv_di_nc <- ivreg(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state) | 
                    valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                    as.factor(year)*as.factor(state), data = q1)

## getting slightly off coefficients
coeftest(iv_ssi_c, vcov = vcovHC(iv_ssi_c, type = "HC1")) %>%
  tidy() 

coeftest(iv_di_c, vcov = vcovHC(iv_di_c, type = "HC1")) %>%
  tidy() # pretty good estimate, but std error is still off

coeftest(iv_ssi_nc, vcov = vcovHC(iv_ssi_nc, type = "HC1")) %>%
  tidy() 

coeftest(iv_di_nc, vcov = vcovHC(iv_di_nc, type = "HC1")) %>%
  tidy() 


```


## Question 6d(iii).D: Summary table

```{r}
## DO WE NEED TO HAVE STANDARD ERRORS AS PARENTHESES OR CAN WE HAVE IT AS A SEPARATE COLUMN?

#lots of kable & bind_rows. need to mutate a column w/ descriptor info on which regression (e.g. ols, control/nc) the coeff corresponds to

sum_table <- function(reg_model, reg_name) {
  coeftest(reg_model, vcov = vcovHC(reg_model, type = "HC1")) %>%
    tidy()  %>%
    mutate(reg = reg_name) %>%
    filter(term == "diff_earnPOW") %>%
    select(reg, estimate, std.error, p.value) %>%
    mutate(
      sig_1perc = if_else(p.value < 0.01, "yes", "no"),
      sig_5perc = if_else(p.value < 0.05, "yes", "no"),
      sig_10perc = if_else(p.value < 0.1, "yes", "no"))
}

q6_sum <- sum_table(ols_di_nc, "OLS Disability - No Control") %>%
  bind_rows(sum_table(ols_di_c, "OLS Disability - Control")) %>%
  bind_rows(sum_table(ols_ssi_nc, "OLS SSI - No Control")) %>%
  bind_rows(sum_table(ols_ssi_c, "OLS SSI - Control")) %>%
  bind_rows(sum_table(iv_di_nc, "IV Disability - No Control")) %>%
  bind_rows(sum_table(iv_di_c, "IV Disability - Control")) %>%
  bind_rows(sum_table(iv_ssi_nc, "IV SSI - No Control")) %>%
  bind_rows(sum_table(iv_ssi_c, "IV SSI - Control"))

q6_sum %>%
  knitr::kable(
    caption = "Question 6d: Summary Table of OLS & IV Regressions", 
    col.names = c(
      "Regression type",
      "Main estimate (log diff in county earnings)", 
      "Robust Standard Error",
      "P-value",
      "Sig @ 1%?",
      "Sig @ 5%?",
      "Sig @ 10%?"),
    digits = 4)


# manipulate manually after exporting to Excel

# bind_rows(sum_table(iv, ""))

#MAYBE
#stargazer(iv_di_nc, iv_di_c, title="Results", align=TRUE)



```

