---
title: "Inequality PSet 1 - Question #6d (code part)"
author: "Julia Du"
date: "`r lubridate::today()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache = TRUE)
```

## Load necessary libraries
```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(stats)
library(ivreg)
library(lmtest)
library(sandwich)
library(moderndive)
library(kableExtra)
library(tinytex)

#didn't end up using library(stargazer) or library(lfe)

theme_set(theme_minimal())
```

## Question 6d(i): Construct the outcome variables

```{r}
q1 <- read_csv("./DataExercise_IV/pset_iv_data.csv") %>%
  rename(pop = Pop, earnPOW = EarnPOW, man = Manufact)

# 6d(i).A: transform var by taking nat logs
q1 <- q1 %>%
  mutate(lnpay = log(pay),
         lnpop = log(pop),
         lnearnPOW = log(earnPOW), # county earnings
         lndSSI = log(dSSI)) # SSI payments 

# 6d(i).B: construct log diff
q1 <- q1 %>%
  group_by(fips) %>% 
  mutate(lagpay = lag(lnpay, n = 1),
         lagpop = lag(lnpop, n = 1),
         lagearnPOW = lag(lnearnPOW, n = 1),
         lagdSSI = lag(lndSSI, n = 1)) %>%
  mutate(diff_pay = lnpay - lagpay,
         diff_pop = lnpop - lagpop,
         diff_earnPOW = lnearnPOW - lagearnPOW,
         diff_dSSI = lndSSI - lagdSSI) 
  
# 6d(i).C: construct var that is 
# fraction of county earnings from manufacturing in 1969
q1 <- 
  q1 %>%
  filter(year == 1969) %>%
  mutate(frac_earn = man/earnPOW) %>%
  select(frac_earn, fips) %>%
  left_join(q1, by = "fips")
```

## Question 6d(ii): Construct the instruments

```{r}
# construct coal price
q1 <- q1 %>%
  mutate(coalprice = pcoalyr/pyr)

# construct change in coal reserves
q1 <- q1 %>%
  group_by(fips) %>%
  mutate(ln_coalprice = log(coalprice),
         lag_coalprice = lag(ln_coalprice, n = 1),
         diff_coalprice = ln_coalprice - lag_coalprice) %>%
  mutate(ln_coalres = log(coalres)) %>%
  mutate(valchange_cr = 
           if_else(coalres > 0,
                   diff_coalprice * ln_coalres, 0)) %>%
  mutate(valchange_crlag1 = lag(valchange_cr, n = 1),
         valchange_crlag2 = lag(valchange_cr, n = 2))
```


## Question 6d(iii).A: OLS structural eqns

```{r}
# construct OLS regressions
ols_ssi_c <- lm(diff_dSSI ~ diff_earnPOW + as.factor(year)*as.factor(state) + 
                  msa + lnpop + diff_pop + frac_earn, data = q1)

ols_ssi_nc <- lm(diff_dSSI ~ diff_earnPOW + 
                   as.factor(year)*as.factor(state), data = q1)

ols_di_c <- lm(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state) + 
                 msa + lnpop + diff_pop + frac_earn, data = q1)

ols_di_nc <- lm(diff_pay ~ diff_earnPOW + 
                  as.factor(year)*as.factor(state), data = q1)

# find estimates & robust SEs
coeftest(ols_ssi_c, vcov = vcovHC(ols_ssi_c, type = "HC1")) %>%
  broom::tidy() 

coeftest(ols_ssi_nc, vcov = vcovHC(ols_ssi_nc, type = "HC1")) %>%
  tidy()

coeftest(ols_di_c, vcov = vcovHC(ols_di_c, type = "HC1")) %>%
  tidy()

coeftest(ols_di_nc, vcov = vcovHC(ols_di_nc, type = "HC1")) %>%
  tidy()
```

NEED TO INTERPRET MAIN ESTIMATES (that's just the estimated coeff for diff_earnPOW aka log diff in county earnings, right?)

## Question 6d(iii).B: 1st-stage estimates

```{r}
first_c <- lm(diff_earnPOW ~ valchange_cr + as.factor(state)*as.factor(year) + 
                msa + lnpop + diff_pop + frac_earn, data = q1)

first_nc <- lm(diff_earnPOW ~ valchange_cr + as.factor(state)*as.factor(year), 
               data = q1)

#get estimates & normal SE (not robust)
get_regression_table(first_c) 
get_regression_table(first_nc) 
```

interpret the main estimates

## Question 6d(iii).C: 2SLS structural eqn

```{r}
# helpful links:
#https://cran.r-project.org/web/packages/ivreg/ivreg.pdf
#https://www.econometrics-with-r.org/12-2-TGIVRM.html

iv_ssi_c <- ivreg(diff_dSSI ~ 
                    diff_earnPOW + as.factor(year)*as.factor(state) + msa + 
                    lnpop + diff_pop + frac_earn 
                  | valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                    as.factor(year)*as.factor(state) + msa + 
                    lnpop + diff_pop + frac_earn, data = q1)


iv_ssi_nc <- ivreg(diff_dSSI ~ diff_earnPOW + as.factor(year)*as.factor(state) 
                   | valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                     as.factor(year)*as.factor(state), data = q1)

iv_di_c <- ivreg(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state) + 
                   msa + lnpop + diff_pop + frac_earn 
                 | valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                   as.factor(year)*as.factor(state) + msa + 
                   lnpop + diff_pop + frac_earn, data = q1)

iv_di_nc <- ivreg(diff_pay ~ diff_earnPOW + as.factor(year)*as.factor(state) | 
                    valchange_cr + valchange_crlag1 + valchange_crlag2 + 
                    as.factor(year)*as.factor(state), data = q1)

# obtain robust SEs
coeftest(iv_ssi_c, vcov = vcovHC(iv_ssi_c, type = "HC1")) %>%
  tidy() 

coeftest(iv_di_c, vcov = vcovHC(iv_di_c, type = "HC1")) %>%
  tidy() 

coeftest(iv_ssi_nc, vcov = vcovHC(iv_ssi_nc, type = "HC1")) %>%
  tidy() 

coeftest(iv_di_nc, vcov = vcovHC(iv_di_nc, type = "HC1")) %>%
  tidy() 
```


## Question 6d(iii).D: Summary table

As discussed with Gin in office hours, I'm writing my robust standard errors for the main estimates in another column, rather than including them in parentheses in the same cell as the estimates.

```{r}
sum_table <- function(reg_model, reg_name) {
  coeftest(reg_model, vcov = vcovHC(reg_model, type = "HC1")) %>%
    tidy()  %>%
    mutate(reg = reg_name) %>%
    filter(term == "diff_earnPOW") %>%
    select(reg, estimate, std.error, p.value) %>%
    mutate(
      sig_1perc = if_else(p.value < 0.01, "yes", "no"),
      sig_5perc = if_else(p.value < 0.05, "yes", "no"),
      sig_10perc = if_else(p.value < 0.1, "yes", "no"))
}

q6_sum <- sum_table(ols_di_nc, "OLS Disability - No Control") %>%
  bind_rows(sum_table(ols_di_c, "OLS Disability - Control")) %>%
  bind_rows(sum_table(ols_ssi_nc, "OLS SSI - No Control")) %>%
  bind_rows(sum_table(ols_ssi_c, "OLS SSI - Control")) %>%
  bind_rows(sum_table(iv_di_nc, "IV Disability - No Control")) %>%
  bind_rows(sum_table(iv_di_c, "IV Disability - Control")) %>%
  bind_rows(sum_table(iv_ssi_nc, "IV SSI - No Control")) %>%
  bind_rows(sum_table(iv_ssi_c, "IV SSI - Control"))

q6_sum %>%
  knitr::kable(
    caption = "Summary Table of OLS and IV Regressions", 
    col.names = c("Regression type",
                  "Main estimate (log diff in county earnings)", 
                  "Robust standard error",
                  "P-value",
                  "Sig @ 1%?",
                  "Sig @ 5%?",
                  "Sig @ 10%?"),
    digits = 5,
    linesep = "\\addlinespace") %>%
  kableExtra::column_spec(2:4, width = "2cm") %>%
  kableExtra::column_spec(5:7, width = "1cm") 
```

